---
title: "Louisiana profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.
```{r}
options(scipen=999)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggrepel)
library(sf)
library(tigris)
library(tidycensus)
census_api_key("c84fed421923b886b2bbdab82b6838c9a11b8c44")
```
## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

**A1.** There are 217334 PPP loan applications

```{r}
la_ppp <- read_csv("louisiana_ppp.csv.zip") %>%
  clean_names()

#how many applications: 217334 minus 2 that are the wrong zip codes, so 217,332.

total_apps <- la_ppp %>%
  summarise(count=n())

#any loans with the same address?

address_la <-la_ppp %>%
  group_by(address) %>% summarise(total=n()) %>% arrange(desc(total))
#yes, tons. 8710 JEFFERSON HIGHWAY has 76 loans at that address apparently. 

jefferson_hwy <- la_ppp %>%
  filter(address == "8710 JEFFERSON HIGHWAY")
# all of these are some sort of "Valman" or "ValCo" company out of Baton Rouge I googled the address and it is the physical building for the corporate address of "Valluzzo Companies."

#sort by zip to see which zips have the highest average ppp loan amount and most loans
la_ppp_zip <- la_ppp %>%
    mutate(zip5 = str_sub(zip, start=1L, end=5L)) %>%
    group_by(zip5) %>%
    summarize(count=n()) %>%
    arrange(desc(count))
# 70508 (Lafayette, LA) had the highest number of ppp loans with 3068 loans, followed by 70301 (Thibodaux, LA) and 70816 (Baton Rouge, LA) which were tied for second with 2968 loans. Looking at the counties with the LEAST amount of loans, I found at least one error. The data included "NA," and also 40503, which is a Kentucky zip code. Another error was 70016 - which is an Italian zip code. 

#Next, I wanted to see if the zip codes with the most loans were also the zip codes with the highest dollar amounts
la_ppp_amount_zip <- la_ppp %>%
    mutate(zip5 = str_sub(zip, start=1L, end=5L)) %>%
    group_by(zip5) %>%
    summarise(
    total_loans = sum(amount),
    count=n()) %>%
  arrange(desc(total_loans))
#Interestingly, the zip code with the highest number of loans did not have the highest dollar amount of loans. 70508 was 3rd with $302,644,379 in total loans. The highest amount of loans was in 70130 (New Orleans) with $396,902,184. 

# Examine NAICS code 
naics_ppp <- la_ppp %>%
  group_by(naics_code) %>%
  summarize(count=n()) %>%
    arrange(desc(count))
#Naics code 812112 had the most loans with 14,578 loans. This code is for beauty salons. This code had nearly ten thousand more loans than any other. The second most common naics code was 541110 with 4,659 loans. This code is for offices of lawyers. The third most common loan was 722511, which is full-service restaraunts, that had 4631 loans. There were 58 naics codes with only 1 loan, including 111332 which is grape vinyards, 212113, which is anthracite mining, and 532220, which is formal wear and costume rental. I actually think looking at the least-common naics codes is more interesting. Naics code 921140 means "This industry comprises government establishments serving as councils and boards of commissioners or supervisors and such bodies where the chief executive (e.g., county executive or city mayor) is a member of the legislative body (e.g., county or city council) itself." This is odd, because woulnd't such people work for the government? 

#Also looking at jobs retained

jobs_retained <- la_ppp %>%
  group_by(jobs_retained) %>%
  summarize(count=n()) %>%
    arrange(desc(count))

# Most loans only retained one job, the second most loans retained two jobs, and so on until 10 and 9, which were flipped. Interestingly, there was 1 loan that retained 0 jobs but received $7,915. 
  
 
```
## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure. Describe the most interesting or newsworthy findings based on your exploration.

**A2.** 
```{r}


```
## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below.

**A3.** 
```{r}
#most approved applications
approved_loans <- la_ppp %>%
  mutate(zip5 = str_sub(zip, start=1L, end=5L)) %>%
  group_by(lender) %>%
  summarize(count=n(), total_amount=sum(amount)) %>%
  mutate(average_amount = total_amount/count) %>%
  arrange(desc(count))

         
         # Prestamos CDFI, LLC is the lender with the most approved loans, with 21,251 and an average loan of $16,560. Capital Plus Financial, LLC had the second most with 16,110 approved loans at an average of $16,564. Neither Prestamos nor Capital Plus Financial had the highest amount of loans. Hancock Whitney Bank, which had the third most approved loans, had the highest amount of loans with $1,926,757,771 and an average loan of $149,755. They are the only lender with over $1 billion in loans. 

#undisbursed loans
undisbursed_loans <- la_ppp %>%
  mutate(zip5 = str_sub(zip, start=1L, end=5L),
         undisbursed = !is.na(undisbursed_amount) & undisbursed_amount > 0) %>%
  group_by(lender) %>%
  summarize(count=n(), total_amount=sum(amount), total_undisbursed_amount=sum(undisbursed_amount),
            undisbursed_n = sum(undisbursed))

la_ppp_totals <- la_ppp %>%
  group_by(lender, undisbursed_amount) %>%
  summarize(n = n())

#For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below.

undisbursed_lenders <- undisbursed_loans %>%
  filter(total_undisbursed_amount > 0)

write_csv(undisbursed_lenders, "undisbursed_lenders.csv")
```

## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** 
```{r}

```
## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.**
```{r}

```