---
title: "Reverse Engineering Project"
author: "Eve Sampson, Maddy Peek, David Andrews"
date: "Mar 5, 2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women shot by police](Shttps://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <-read_csv("data/fatal-police-shootings-data.csv")



```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Since 2015, police have fatally shot nearly 250 women. Like Taylor, 89 of them were killed at homes or residences where they sometimes stayed.”

* **Analysis summary**: We were able to confirm the first part of the sentence; by filtering for gender and dates before publication, we got 248 fatalities. We are unable to discern if they were killed at their homes or residences where they stayed based on the data. We would need to check the locations against their home addresses to verify. 

```{r}
  women_shot_since_2015 <- fatal_police_shootings %>% mutate(year = year(date)) %>% filter(gender == "F", date < "2020-09-04")

#this gives us 248 fatalities.
  #women_shot_since_2015 <- fatal_police_shootings %>% mutate(year = year(date)) %>% filter(gender == "F") summarize(count = n()) this gives you 320 but includes data after publication




```

### Sentence 2

* **Sentence text**: “The Post began tracking fatal police shootings in a database in January 2015, months after a White police officer in Ferguson, Mo., killed Michael Brown, an unarmed Black man during a confrontation. Since then, police have shot and killed about 1,000 people a year.”
* **Analysis summary**: Yes. Grouping the dates by month and then again by year, there have been about 1,000 people fatally shot by police every year since 2015.

```{r}

fatal_police_shootings_data_year <- fatal_police_shootings %>% mutate(year = year(date))


fatal_police_shootings_data_year_organized <-fatal_police_shootings_data_year %>% 
group_by(year) %>%
 summarize(
    count = n())


```

```{r}



```

### Sentence 3

* **Sentence text**: “Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}



```

### Sentence 4

* **Sentence text**: “About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”



* **Analysis summary**: We came pretty close. Instead of 77 of the 247 women noted in the sentence, we got 79/248 filtering for the conditions outlined in the sentence. We did, however, get very similar percentages, successfully illustrating the point of the sentence. We think there are a few unknowns that account for these slight variations in data. 1. We don't know if additional data was entered for this timeframe after the article was published. Perhaps more cases were added/more conditions were included that became known after the published date. 2. We don't know what specific date was used as the cutoff date when this article was published. Since it was published on 9/4/20, we figured any entries that occured on 9/4/20 itself was likely not included in the statistics. Playing around with the cutoff date as early as 9/1/20 didn't produce exact results either. Still, we were able to reproduce the overarching point of the sentence, which is that mental health was a resulting factor in about 31% of women fatally shot by police since 2015, while mental health only accounted for about 22% of the men fatally shot by police since 2015.

```{r}

fatal_police_shootings_before_article <- fatal_police_shootings_data_year %>%
  filter(date < "2020-09-04")
#this gets us all entries that happened before 2020-09-04 when the article was published.

women_shot_mentalhealth <- fatal_police_shootings_before_article %>%
filter(gender == 'F') %>% 
#nrow() = 248 rows
filter(signs_of_mental_illness == 'TRUE') %>% 
  filter(manner_of_death == "shot") #%>% 
 # nrow() = 79
#(79/248)*100 = 31.8%

#79 of the 248 of women who were fatally shot by police, or 31.8%, were reported to show signs of mental illness.

men_shot_mentalhealth <- fatal_police_shootings_before_article %>%
filter(gender == 'M') %>% 
#nrow() = 5362
filter(signs_of_mental_illness == 'TRUE') %>% 
filter(manner_of_death == "shot") #%>% 
#nrow() = 1,153
#(1153/5362)*100 = 21.5%

#21.5% of men who were fatally shot by police were reported to show signs of mental illness.


```

### Sentence 5

* **Sentence text**:“The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings. Of those fatally shot every year, about 44 have been women.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}


```

-30-